---
title: "A Comparison of the Exponential Distribution with the Central Limit Theorem"
author: "Ian Dobbs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE, fig.align = "center",
                      fig.path="figures/", options(scipen=999, digits = 2))
knitr::opts_current$get('label')
```
```{r captioner, message=FALSE, echo=FALSE}
# use captioner to add figure number and caption
library(captioner)
fig_nums <- captioner()
fig_nums("figa", "A table")
fig_nums("figb", "A table")
```
## Synopsis

In this report I aim to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The report will illustrate the properties of the distribution via the simulation and explanation of the distribution of the mean of 40 exponentials.

## Simulations

text
```{r simulations, cache=TRUE}

```

## Sample Mean versus Theoretical Mean

text
```{r smean, cache=TRUE}

```

## Sample Variance versus Theoretical Variance

text
```{r tmean}

```

## Distribution

text
```{r distribution, message=FALSE}

```

## Conclusions

text
