---
title: "A Comparison of the Exponential Distribution with the Central Limit Theorem"
author: "Ian Dobbs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE, fig.align = "center",
                      fig.path="figures/", options(scipen=999))
knitr::opts_current$get('label')
```
```{r captioner, message=FALSE, echo=FALSE}
# use captioner to add figure number and caption
library(captioner)
fig_nums <- captioner()
fig_nums("figa", "A histogram of the distribution of averages of 40 exponentials over 1000 simulations. Fitted with a normal curve")
fig_nums("figb", "A table")
```
## Synopsis

In this report I aim to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The report will illustrate the properties of the distribution via the simulation and explanation of the distribution of the mean of 40 exponentials.

## Simulations

This report uses the rexp(n, lambda) function in R to calculate the mean of a random sample of 40 exponentials and repeat for one thousand simulations. The sample mean and variance of the resulting distribution of averages will be compared to the theoretical mean and variance.
```{r simulations}
lambda <- 0.2 # the rate parameter of rexp function
mu <- 1/lambda # mean of the exponential distribution
sd <- 1/lambda # standard deviation of the exponential distribution
n <- 40 # number of sample exponentials
se <- sd / sqrt(n) # 
svar <- sd^2 / n # 
nosim <- 1000 # number of simulations

set.seed(1) # to make the result reproducible
x <- NULL
for (i in 1:nosim) x = c(x, mean(rexp(n, lambda))) # create the distribution of a large collection (1000)
# of averages of 40 # exponentials
```

## Sample Mean versus Theoretical Mean

 The sample mean and the theoretical mean are both measures of central tendency. Comparing the sample mean to the theoretical mean demonstrates that the sample mean is an estimate of the population mean.
```{r smean}
round(c(mean(x), mu), 2)
```

## Sample Variance versus Theoretical Variance

text
```{r tmean}

```

## Distribution

text
```{r distribution, fig.cap=fig_nums("figa")}
# create the X-axis grid to overlay the normal curve
x2 <- seq(min(x), max(x), length = length(x))

# create the Normal curve
fun <- dnorm(x2, mean = mean(x), sd = sd(x))

# create the Histogram of the simulation
par(mfrow = c(1,2))
hist(x, prob = TRUE, # probability = true produces a density plot 
     ylim = c(0, max(fun)),
     main = "", xlab = "")
lines(x2, fun, col = 2, lwd = 2)
abline(v = mean(x), col="red", lwd=3, lty=2)
abline(v = mean(mu), col="blue", lwd=3, lty=5)

set.seed(1)
y <- rexp(nosim, lambda) # create the distribution of a large collection (1000) of random exponentials
hist(y, prob = TRUE, freq = FALSE, main = "", xlab = "")
curve(dexp(x, rate=0.2, log=FALSE), add = TRUE)
```

## Conclusions

text
