---
title: "A Comparison of the Exponential Distribution with the Central Limit Theorem"
author: "Ian Dobbs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE, fig.align = "center",
                      fig.path="figures/", options(scipen=999, digits = 2))
knitr::opts_current$get('label')
```
```{r captioner, message=FALSE, echo=FALSE}
# use captioner to add figure number and caption
library(captioner)
fig_nums <- captioner()
fig_nums("figa", "A table")
fig_nums("figb", "A table")
```
## Synopsis

In this report I aim to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The report will illustrate the properties of the distribution via the simulation and explanation of the distribution of the mean of 40 exponentials.

## Simulations

text
```{r simulations}
lambda <- 0.2 # the rate parameter of rexp function
mu <- 1/lambda # mean of the exponential distribution
sd <- 1/lambda # standard deviation of the exponential distribution
n <- 40 # number of exponentials
se <- sd / sqrt(n) # 
svar <- sd^2 / n # 
nosim <- 1000 # number of simulations

set.seed(1) # to make the result reproducible
x <- NULL
for (i in 1:nosim) x = c(x, mean(rexp(n, lambda)))

# create the X-axis grid to overlay the normal curve
x2 <- seq(min(x), max(x), length = length(x))

# create the Normal curve
fun <- dnorm(x2, mean = mean(x), sd = sd(x))

# create the Histogram of the simulation
hist(x, prob = TRUE, # probability = true produces a density plot 
     ylim = c(0, max(fun)),
     main = "Histogram with normal curve")
lines(x2, fun, col = 2, lwd = 2) 
```

## Sample Mean versus Theoretical Mean

text
```{r smean}

```

## Sample Variance versus Theoretical Variance

text
```{r tmean}

```

## Distribution

text
```{r distribution}

```

## Conclusions

text
