---
title: "Basic Inferential Data Analysis - The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Ian Dobbs"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE, fig.align = "center",
                      fig.path="figures/", options(scipen=999))
knitr::opts_current$get('label')
```
```{r captioner, message=FALSE, echo=FALSE, results='hide'}
# use captioner to add figure number and caption
library(captioner)
fig_nums <- captioner()
fig_nums("figa", "A scatter plot of the ToothGrowth data: length vs dose, given type of supplement. The red triangles representing the average length at each dose for the given type of supplement.")
```
## Synopsis

The ToothGrowth dataset shows the effect of vitamin C on tooth growth in guinea pigs. The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

## Summary of the data

Text here

```{r ToothGrowth}
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
```

Text here

```{r toothgrowth, echo=FALSE, fig.cap=fig_nums("figa")}
library(ggplot2)
ggplot(ToothGrowth, aes(x=factor(dose), y=len, fill=factor(dose), color=factor(dose)))+
        geom_point()+
        stat_summary(geom = "point",
                     fun = "mean",
                     col = "black",
                     size = 3,
                     shape = 24,
                     fill = "red")+
        facet_grid(.~supp)+
        labs(x = "Dose", y = "Length")
```

## Hypothesis tests

## Conclusions

## Assumptions